---
title: "Water Usage Statistics"
output: ioslides_presentation
---

```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)

source("Utilities.R")

# Read from formatted file
# Main Heater numbers per housing unit
fuel_water_units <- read_excel(
  sheet=1,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterUnits.xls",
  na=c("Q","N"))
size_water_units <- read_excel(
  sheet=2,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterUnits.xls",
  na=c("Q","N"))
age_water_units <- read_excel(
  sheet=3,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterUnits.xls",
  na=c("Q","N"))

fuel_water_members <- read_excel(
  sheet=2,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterMembers.xls",
  na=c("Q","N"))
size_water_members <- read_excel(
  sheet=3,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterMembers.xls",
  na=c("Q","N"))
age_water_members <- read_excel(
  sheet=1,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterMembers.xls",
  na=c("Q","N"))

fuel_water_years <- read_excel(
  sheet=2,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterYears.xls",
  na=c("Q","N"))
size_water_years <- read_excel(
  sheet=1,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterYears.xls",
  na=c("Q","N"))
age_water_years <- read_excel(
  sheet=3,
  "~/Documents/Dev/DataScience/HomeWaterHeater/hc8_all/HomeWaterYears.xls",
  na=c("Q","N"))
```


## Examples of Water & Energy waste

- "The average water heater for a family of 5+ is around 50 gallons"
- Number of household members * 12 = about average number of gallons used
- The truth in the U.S. is inverted.

## Number of Household Members vs Count by Heater Size

```{r membersVsSize}
exploreTypeVsCount(
  tidyData(size_water_members),
  "Size of main water heater",
  "Size"
  )
```

## Even more explicit

```{r warning=F}
# Average household need # of members * 12 gallons of water
swm_tidy <- tidyData(size_water_members[1:3,]) %>%
  mutate("Average Waste" = rep(c(30,49,60),5) - sort(rep(1:5,3)) * 12) %>%
  mutate("Wasting water" = `Average Waste` < 0)

ggplot(data=swm_tidy,aes(x=`Unit Type`,fill=`Size of main water heater`)) +
  geom_col(aes(y=`Unit Count (million)`),position="dodge") +
  geom_point(aes(y=`Average Waste`,size=`Average Waste`),position=position_dodge(width = 0.9)) +
  guides(fill=guide_legend(title="Size of heater")) +
  xlab("") + ggtitle("Heater Size and Average Waste versus # of Household Members") + 
  scale_y_continuous(sec.axis = sec_axis( ~ .,name = "Gallons of waste"))

```

## Water Heater Type in Households based on Year of Construction

```{r fuelWaterYears}
exploreTypeVsCount(
  tidyData(fuel_water_years),
  "Fuel used by main water heater",
  "Fuel source"
  )
```

## Same, but based on % to demonstrate progression

```{r echo=F,warning=F,message=F}
fwy_tidy <- tidyData(fuel_water_years)
fwy_freq <- group_by(
  fwy_tidy,
  "Unit Type"=get(colnames(fwy_tidy)[2])) %>%
  summarize("Total"=sum(`Unit Count (million)`))

fwy_ratio <- left_join(fwy_tidy,fwy_freq) %>%
  mutate("Percentage"=`Unit Count (million)`/Total * 100)

ggplot(data=fwy_ratio,aes(x=`Unit Type`,y=`Percentage`,
                             fill=`Fuel used by main water heater`)) +
  geom_col() +
  guides(fill=guide_legend(title="Fuel source")) +
  xlab("")
```


